# 1. Fourier Transform
  - reference: <http://www.di.fc.ul.pt/~jpn/r/fourier/fourier.html>
    + 2013.  
    
## 1.1 Complex Wave

```{r, echo = FALSE}
setwd("~/R/workspace/plant/script/exercise")

xs <- seq(-2*pi,2*pi,pi/100)
wave.1 <- sin(3*xs)
wave.2 <- sin(10*xs)
par(mfrow = c(1, 2))
plot(xs,wave.1,type="l",ylim=c(-1,1)); abline(h=0,lty=3)
plot(xs,wave.2,type="l",ylim=c(-1,1)); abline(h=0,lty=3)
```

```{r, echo = FALSE}
wave.3 <- 0.5 * wave.1 + 0.25 * wave.2
plot(xs,wave.3,type="l"); title("Eg complex wave"); abline(h=0,lty=3)
```

## 1.2 Fourier Series
  - Joseph Fourier showed that any periodic wave can be represented by a sum of simple sine waves.  
```{r, echo = FALSE}
wave.4 <- wave.3
wave.4[wave.3>0.5] <- 0.5
plot(xs,wave.4,type="l",ylim=c(-1.25,1.25)); title("overflowed, non-linear complex wave"); abline(h=0,lty=3)
```

```{r, echo = FALSE}
repeat.xs     <- seq(-2*pi,0,pi/100)
wave.3.repeat <- 0.5*sin(3*repeat.xs) + 0.25*sin(10*repeat.xs)
plot(xs,wave.3,type="l"); title("Repeating pattern")
points(repeat.xs,wave.3.repeat,type="l",col="red"); abline(h=0,v=c(-2*pi,0),lty=3)
```


```{r, echo = FALSE}
plot.fourier <- function(fourier.series, f.0, ts) {
  w <- 2*pi*f.0
  trajectory <- sapply(ts, function(t) fourier.series(t,w))
  plot(ts, trajectory, type="l", xlab="time", ylab="f(t)"); abline(h=0,lty=3)
}

# An eg
plot.fourier(function(t,w) {sin(w*t)}, 1, ts=seq(0,1,1/100)) 
```

```{r, echo = FALSE}
acq.freq <- 100                    # data acquisition frequency (Hz)
time     <- 6                      # measuring time interval (seconds)
ts       <- seq(0,time,1/acq.freq) # vector of sampling time-points (s) 
f.0      <- 1/time                 # fundamental frequency (Hz)

dc.component       <- 0
component.freqs    <- c(3,10)      # frequency of signal components (Hz)
component.delay    <- c(0,0)       # delay of signal components (radians)
component.strength <- c(.5,.25)    # strength of signal components

f <- function(t,w) { 
  dc.component + 
  sum( component.strength * sin(component.freqs*w*t + component.delay)) 
}

plot.fourier(f,f.0,ts)   
```

## 1.3 Phase Shifts

```{r, echo = FALSE}
component.delay <- c(pi/2,0)       # delay of signal components (radians)
plot.fourier(f,f.0,ts)   
```

## 1.4 DC Components

```{r, echo = FALSE}
dc.component <- -2
plot.fourier(f,f.0,ts)  
```

## 1.5 Fourier Transform
  - a stationary trajectory:
```{r, echo = FALSE}
library(stats)
fft(c(1,1,1,1)) / 4  # to normalize
```

  - an increasing trajectory:
```{r, echo = FALSE}
fft(1:4) / 4
```

  - a little function to convert the fft() output to the animation output:

```{r, echo = FALSE}
# cs is the vector of complex points to convert
convert.fft <- function(cs, sample.rate=1) {
  cs <- cs / length(cs) # normalize

  distance.center <- function(c)signif( Mod(c),        4)
  angle           <- function(c)signif( 180*Arg(c)/pi, 3)
  
  df <- data.frame(cycle    = 0:(length(cs)-1),
                   freq     = 0:(length(cs)-1) * sample.rate / length(cs),
                   strength = sapply(cs, distance.center),
                   delay    = sapply(cs, angle))
  df
}

convert.fft(fft(1:4))
```

  - here’s a function that applies the previous equation, ie, makes the IFT:
  
```{r, echo = FALSE}
# returns the x.n time series for a given time sequence (ts) and
# a vector with the amount of frequencies k in the signal (X.k)
get.trajectory <- function(X.k,ts,acq.freq) {
  
  N   <- length(ts)
  i   <- complex(real = 0, imaginary = 1)
  x.n <- rep(0,N)           # create vector to keep the trajectory
  ks  <- 0:(length(X.k)-1)
  
  for(n in 0:(N-1)) {       # compute each time point x_n based on freqs X.k
    x.n[n+1] <- sum(X.k * exp(i*2*pi*ks*n/N)) / N
  }
  
  x.n * acq.freq 
}
```

  - plot.frequency.spectrum() plot a frequency spectrum
  - plot.harmonic() plots the i-th harmonic on the current plot

```{r, echo = FALSE}
plot.frequency.spectrum <- function(X.k, xlimits=c(0,length(X.k))) {
  plot.data  <- cbind(0:(length(X.k)-1), Mod(X.k))

  # TODO: why this scaling is necessary?
  plot.data[2:length(X.k),2] <- 2*plot.data[2:length(X.k),2] 
  
  plot(plot.data, t="h", lwd=2, main="", 
       xlab="Frequency (Hz)", ylab="Strength", 
       xlim=xlimits, ylim=c(0,max(Mod(plot.data[,2]))))
}

# Plot the i-th harmonic
# Xk: the frequencies computed by the FFt
#  i: which harmonic
# ts: the sampling time points
# acq.freq: the acquisition rate
plot.harmonic <- function(Xk, i, ts, acq.freq, color="red") {
  Xk.h <- rep(0,length(Xk))
  Xk.h[i+1] <- Xk[i+1] # i-th harmonic
  harmonic.trajectory <- get.trajectory(Xk.h, ts, acq.freq=acq.freq)
  points(ts, harmonic.trajectory, type="l", col=color)
}
```

```{r, echo = FALSE, collapse=TRUE}
X.k <- fft(c(4,0,0,0))                   # get amount of each frequency k

time     <- 4                            # measuring time interval (seconds)
acq.freq <- 100                          # data acquisition frequency (Hz)
ts  <- seq(0,time-1/acq.freq,1/acq.freq) # vector of sampling time-points (s) 

x.n <- get.trajectory(X.k,ts,acq.freq)   # create time wave

plot(ts,x.n,type="l",ylim=c(-2,4),lwd=2)
abline(v=0:time,h=-2:4,lty=3); abline(h=0)

plot.harmonic(X.k,1,ts,acq.freq,"red")
plot.harmonic(X.k,2,ts,acq.freq,"green")
plot.harmonic(X.k,3,ts,acq.freq,"blue")
```


## 1.6.Examples

```{r, echo = FALSE}
acq.freq <- 100                    # data acquisition (sample) frequency (Hz)
time     <- 6                      # measuring time interval (seconds)
ts       <- seq(0, time-1/acq.freq, 1/acq.freq) # vector of sampling time-points (s) 
f.0 <- 1/time

dc.component <- 1
component.freqs <- c(3,7,10)        # frequency of signal components (Hz)
component.delay <- c(0,0,0)         # delay of signal components (radians)
component.strength <- c(1.5,.5,.75) # strength of signal components

f   <- function(t,w) { 
  dc.component + 
  sum( component.strength * sin(component.freqs*w*t + component.delay)) 
}

plot.fourier(f,f.0,ts=ts)
```

```{r, echo = FALSE}
w <- 2*pi*f.0
trajectory <- sapply(ts, function(t) f(t,w))
head(trajectory, n=30)
```

```{r, echo = FALSE}
X.k <- fft(trajectory)                   # find all harmonics with fft()
plot.frequency.spectrum(X.k, xlimits=c(0,acq.freq/2))
```


```{r}
plot(abs(X.k[1:300])^2, type="l")
```

```{r, echo = FALSE}
spectrum(trajectory, method = "pgram", taper=0, fast=FALSE, detrend=FALSE, log="no")
```


  - testing
```{r, echo = FALSE}
spectrum(trajectory)
```

```{r, echo = FALSE}
spectrum(trajectory, method = "ar")
```

```{r, echo = FALSE}
x.n <- get.trajectory(X.k, ts, acq.freq) / acq.freq  # TODO: why the scaling?
plot(ts,x.n, type="l"); abline(h=0,lty=3)
points(ts,trajectory,col="red",type="l") # compare with original
```



